### 2020-06-05 outline.Rmd
rename package

---
title: "Outline"
output: github_document
---

# Introduction
  - MEME Suite is an excellent collection of tools for motif analysis
  - MEME Suite Webserver serves thousands of users/year
  - Problem: Manual uploads & configuration of each job makes iteration difficult
  - Problem: occasionally long queue times
  - Problem: Although data are returned in machine-readable format (.xml), there are few tools which can parse these data, requiring researchers to write their own solutions.
	- Genomics requires high throughput
  - Together, this can increase friction in integrating the MEME suite into an analysis workflow and may result in these powerful tools being underutilized during an analysis.
  - We made {memer}, which integrates the most popular MEME suite tools with the bioconductor stack to allow simple, rapid development of motif analysis steps.
		 - enables more iterations to identify best parameters
		 - describe uses of new functionality & improvied efficiency (briefly)
	- Also novel data visualization
		 - There are currently limited ways to visualize MEME data, but by bringing data into R, opens up the strengths of R data viz.
	** Couch some problems & soutiuons as novel develppments by this package**
	- explicitly say not changing core functions of MEME

# Main Text
  - A key design principle was to create tools that can be integrated in existing analysis pipelines, in addition to allowing users to create new ones.
## Figure: flowchart of each step, getsequence-> run-> import -> visualize -> goto run
## Overarching design principles/{memer} framework
  - MEME utilities work with files on disk & outputs data to files on disk
  - {memer} works by handling writing files from R object inputs, constructing the commandline calls, then reading in the data from the output files.
      - by default, this is all done using temporary files, so the data is not permanently stored on disk.
      - This is to shift the perspective from needing to save the files, to focusing on writing reproducible code which can regenerate the data when run.
      - However, users can set the `outdir` argument to purposefully save the outputs to disk if needed.
        - Also, if users use a 'fasta' file as input, the output data will be stored alongside the fasta file location
  - For functions that use MEME commandline tools, they need the location of 'meme/bin/' on disk, we provide utilities for making this detection simple to reduce friction of using the tool.
    - `check_meme_install` does user-friendly check that meme can be detected from R environment
    - Environment variable `MEME_PATH` (set & forget), `options(meme_path)` (configure for each script), `meme_path` argument, for each function call
  - Whenever possible, {memer} uses standard R data types for better integration with common analysis tools, like `dplyr`, `GenomicRanges`, or base R. In this way, users don't need to learn new ways of working with data.

## Core Utilities
  - We built a common interface to each MEME utility. 
    - Tools taking sequence inputs can use fasta files as input, or a Biostrings representation of sequences.
    - Tools taking motif input can use .meme file path, universalmotif list, or the results object of functions returning motifs as output (allows seamless chaining of tools together)
  - All commandline arguments can be passed to each function, if users mistype the function error handling will suggest typo fixes.
  - `runDreme()`, `runMeme()` = Denovo motif discovery
     - Can be slow
  - `runAme()` = motif enrichment testing
  - `runTomTom()` = motif comparison (often used for matching denovo motifs to known database)
     - both ame & tomtom can use the `meme_db` setting which takes .meme path input, or universalmotif list.
     - can set & forget default in .Renviron
  - `runFimo()` if used with built-in helpers returns genomic coordinates of each motif in input sequences.
     - integrates well with `plyranges` to facilitate range-based join operations to integrate peak-level metadata with motif information, vice-versa
  - for dreme/ame/meme, we've paid special attention to enabling discriminative analysis
     - users can easily perform multiple comparisons of condtion1 vs 2, etc.
     - No special programming is required, no new patterns of data manipulation, etc. Just set input = list_of_sequences, control = "name(s) of entry in list" to compare all entries in list against that set of data (or pooled set if multiple entries).

## Data Visualization
  - `view_tomtom_hits` allows simple comparison of motifs w/ their matches, allowing users to quickly identify & use their domain-specific knowledge to decide if top hit is not "best" hit in context of experiment.
    - Will display "No Match" if there's no hit
    - can use the `force_best_match()` function to alter the best match info
     - see motif data manipulation section for more details
  - `ame_plot_heatmap` allows users to visualize patters of motif enrichment across multiple groups
     - `scale_max` & method = "normalize" provide additional ways to view the data
     - ame_compare_heatmap_methods helper gives users a way to select cutoffs & see tradeoffs of different visualizations, our package documentation (at <website> and in vignettes) go into great detail of how to reason about these differences.
     - function is built in `ggplot2` which can be used to further customize the heatmap
  - These plots can be exported as high-res SVGs for figures in papers

## Motif Metadata manipulation
  - Motif metadata is stored as a data.frame w/ special "motif" column holding the universalmotif object.
  - the `update_motifs()` function uses the data.frame values to update the motif metadata
    - allows users to manipulate motif metadata in bulk, then commit changes to motifs, which can be written to file using universalmotif `write_` functions, or used in plots, more {memer} functions, etc.
    - facilitates manipulation of `MotifDb` metadata, or integrating database information with gene expression, to, for example, filter the search space to only expressed TFs. Code examples of this are described in package documentation.

## Documentation & Tutorials
 - {memer} website & vignettes contain detailed practical tutorials on how to perform comprehensive analysis of ChIP-seq data, query MotifDb to build a database of expressed TFs, and more.

# Conclusion
 - {memer} fits into a broader framework to enable deep analysis of motif content of sequence data.

# Consider adding??
## Most MEME Suite Tools can run fine on a laptop
  - The webserver may give the appearance that lots of computational resources are needed for these tools,
    this isn't really true for TomTom, AME, and Fimo. DREME & MEME Have reasonable runtimes under certain conditions, but can be resource heavy depending on the task.
  - ? Provide data ?
    - Idea would be to run a set of tests on a laptop w/ i5 processor and ~4-8GB RAM (Like a Macbook), 
      and compare to a compute node w/ 16GB RAM, faster CPU. Just to show that you don't need a workhorse for it.