### 2020-11-09 manuscript.Rmd
More edits.

---
title: 'Memes: an R interface to the MEME Suite'
author:
- ORCID: 0000-0003-1000-1579
  name: Spencer L. Nystrom
- ORCID: 0000-0001-8226-0604
  name: Daniel J. McKay
output: 
 - word_document
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
# Figure caption: https://stackoverflow.com/questions/31926623/figures-captions-and-labels-in-knitr
suppressPackageStartupMessages(library(memes))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(GenomicRanges))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ComplexHeatmap))
drake::r_make()
```

# Abstract

Identification of biopolymer motifs represents a key step in the analysis of
biological sequences. The MEME Suite is a widely used toolkit for comprehensive
analysis  of biopolymer motifs; however, these tools are poorly integrated
within popular analysis frameworks like the R/Bioconductor project, creating
barriers to their use. Here we present `memes`, an R package which provides a
seamless R interface to the MEME Suite. `memes` provides a novel "data aware"
interface to these tools, enabling rapid and complex discriminative motif analysis workflows.
In addition to interfacing with popular MEME Suite tools, `memes` leverages
existing R/Bioconductor data structures to store the complex, multidimensional
data returned by MEME Suite tools for rapid data access and manipulation.
Finally, `memes` provides data visualization capabilities to facilitate
communication of results. `memes` is available at
[github.com/snystrom/memes](github.com/snystrom/memes).

# Introduction

Biopolymers, such as DNA and protein, perform varying functions based on their
primary sequence. Short, repeated sequences, or "motifs" represent functional
units within biopolymers that can act as interaction surfaces, create structure,
or have enzymatic activity. Identification of shared motifs across multiple
sequences can provide evidence for shared function, for example, identification
of kinase substrates based on similarities in phosphorylation site sequence, or
characterizing DNA elements based on shared transcription factor binding
sequences (TODO: ref kinase motif:
https://doi.org/10.1016/0968-0004(90)90073-K). Thus the ability to identify,
classify, and compare motifs represents a key step in the analysis of biological
sequences.

The MEME Suite is a widely utilized set of tools to interrogate motif content
across a broad range of biological contexts \cite{meme-suite}. With over 22,000
citations to date, and greater than 30,000 unique users of the webserver
implementation annually, the MEME Suite has emerged as a standard tool in the
field \cite{meme-usage-report-page, google-scholar-meme}. However, several
factors limit the full potential of these tools for use in data analysis. MEME
Suite tools require carefully formatted inputs to achieve full functionality,
yet few tools exist to simplify the process of data formatting, requiring
instead that users write custom code to prepare their data, or prepare the
inputs by hand, both of which have the potential to be error prone without
rigorous testing.  Furthermore, the output data from each MEME Suite tool often
have complex structures that must be parsed to extract the full suite of
information, again requiring users to write custom code for this task. Finally,
although the data-generation capabilities of the MEME Suite are excellent, the
tools lack powerful ways to visualize the results. Collectively, these factors
act as barriers to adoption, preclude deeper analysis of the data, and limit
communication of results to the scientific community.

Here we present `memes`, a motif analysis package that provides a seamless
interface to MEME Suite tools within R. `memes` uses base R and Bioconductor
data types for data input and output, facilitating better integration with
common analysis tools like the `tidyverse` and other Bioconductor packages.
Unlike the commandline implementation, `memes` outputs also function as inputs
to other MEME Suite tools, allowing simple construction of motif analysis
pipelines without additional data processing steps. Additionally, R/Bioconductor
data structures provide a full-featured representation of MEME Suite output
data, providing users quick access to all relevant data structures with simple
syntax. **TODO: Depending on length restrictions, end here**

In addition to wrapping MEME Suite utilities, `memes` extends the functionality
of `universalmotif` to provide base R and `tidyverse` compatible syntax for
manipulating and analyzing motif metadata. These features allow users to
integrate external data with motif databases to aid analysis, such as filtering
a set of transcription factor motifs based on gene expression in the tissue of
interest, then using that filtered set to search for motifs associated with
expressed factors. By maintaining compatability with `universalmotif`, `memes`
outputs can be readily used in a diverse set of tools beyond the `memes`
framework.

`memes` also provides simple utilities for data visualization where appropriate.
Providing R representations of MEME Suite data objects enables generation of
novel data visualizations using the well-established tools available in the R
landscape. Therefore, `memes` empowers users to deeply analyze and visualize
motif analysis data.

`memes` is designed for maximum flexibility and ease of use to allow users to
iterate rapidly during analysis. Here we present several examples of how `memes`
allows novel analyses of ChIP-seq and FAIRE-seq data by facilitating seamless
interoperability between MEME Suite tools and the broader R package landscape.

## Design & Implementation

### Core Utilities

MEME Suite tools are run on the commandline and use files stored on-disk as
input while returning a series of output files containing varying data types. 
As a wrapper of MEME tools, `memes` functions similarly by assigning each
supported MEME Suite tool to a `run` function (`runDreme()`,
`runMeme()`, `runAme()`, `runFimo()`, `runTomTom()`), which
internally writes input objects to files on disk, runs the tool, then imports
the data as R objects. These functions accept sequence and motif inputs as
required by the tool. Sequence inputs are accepted in `BioStrings` format, an
R/Bioconductor package for storing biopolymer sequence data \cite{BioStrings}.
Motif inputs are passed as `universalmotif` objects, another R/Bioconductor
package for representing motif matrices and their associated metadata
\cite{universalmotif}. `memes` `run` functions will also accept paths to files
on disk, such as fasta files for sequence inputs, and meme format files for
motif inputs, reducing the need to read large files into memory. Finally, each
`run` function contains optional function parameters mirroring the commandline
arguments provided by each MEME Suite tool. In this way, `memes` provides a
feature-complete interface to the supported MEME Suite tools.

### Output Data Structures

MEME tools return HTML reports for each tool that display data in a
user-friendly way; however, these files are not ideal for downstream processing.
Depending on the tool, data are also returned in tab-separated, or XML format,
which are more amenable to computational processing. However, in the case that
tab-separated data are returned, results are often incomplete. For example, the
TomTom tool, which compares query motifs to a database of known motifs, returns
tab-separated results that do not contain the frequency matrix representations
of the matched motifs. Instead, users must write custom code to parse these
matrices back out from the input databases, creating additional barriers to
analysis. In the case that data are returned in XML format, these files contain
all relevant data; however, XML files are difficult to parse, again requiring users to
write custom code to extract the necessary data. Finally, the data types
contained in MEME Suite outputs are complex and multidimensional, and thus
require special data structures to properly organize the data in memory.

`memes` provides data import functions for each supported MEME Suite tool, which
import these data as modified R data.frames (described in detail below). These
functions can be called directly by users to import data previously generated by
the commandline or webserver versions of the MEME Suite for use in R. These
import functions also underlie the import step internal to each of the `run`
functions.

### structured data.frames hold multidimensional output data

Data returned from MEME Suite tools are complex and multidimensional, making
them difficult to represent in a simple data structure. For example, MEME and
DREME return de-novo discovered motifs from query sequences along with
statistical information about their enrichment. In this instance, a position
frequency matrix (PFM) of the discovered motif can be represented as a matrix,
while the properties of that matrix (e.g. the name of the motif, the E-value
from the enrichment test, background nucleotide frequency, etc.) must be encoded
outside of the matrix, while maintaining their relationship to the
corresponding PFM. However, other MEME Suite tools, such as TOMTOM, which
compares one motif to a series of several motifs to identify possible matches,
produce an additional layer of complexity such that input matrices (which
contain metadata as previously described) can have a one-to-many relationship
with other motif matrices (again with their own metadata). Thus an ideal
representation for these data is one that can hold an unlimited number of motif
matrices and their metadata, contain complex hierarchies, and be easily
manipulated using standard analysis tools.

The `universalmotif` data structure is a powerful R/Bioconductor representation
for motif matrices and their associated metadata \cite{universalmotif}. However,
`universalmotif` objects are not ideal for performing bulk manipulation or
queries of metadata, nor do they support complex data structures such as nested
elements. R data.frames support nesting arbitrary data types within columns, but
are not sufficient to properly store motif matrices. In order to create an
extensible, nested data structure for tracking MEME Suite data, `memes`
leverages the power of R data.frames and the `universalmotif` format by nesting
`universalmotif` objects within data.frames. These universalmotif data.frames
store information about motifs along rows, where each column stores metadata
associated with a motif, including a `motif` column holding a `universalmotif`
representation of that motif, which can be manipulated by altering columns of
the data.frame (described in detail below). Adding one-to-many relationships within this
structure is trivial, as additional unviersalmotif data.frames can be nested
within eachother. These universalmotif data.frames form the basis for a majority
of `memes` data outputs. These structures are also valid input types to `memes`
functions, and when used as such, output data are appended as new columns to the
input data, ensuring data provenance. Finally, because the universalmotif
data.frame is a standard R data.frame, it can be manipulated using base R or
`tidyverse` functions, making it easier to interface with these
structures.

**Figure?: showing how the data.frame connects to universalmotif**

The universalmotif data.frames generated by memes are not limited to storing
motifs from MEME Suite data. They can also be used to manipulate or store other
descriptive properties of motifs from multiple sources. `universalmotif`
provides several import functions which can convert motifs from external
databases queried from the `MotifDb` package, or from files in a variety of
non-meme file formats \cite{motifdb, universalmotif}. Once imported in `universalmotif` format
the `memes` function `as_universalmotif_dataframe()` converts any
`universalmotif` object into a universalmotif data.frame. Similar to MEME Suite
data stored in this format, if changes are made to the linked metadata columns,
the linked universalmotif object can be updated by calling `update_motifs()` to
retain the data.frame structure for further analysis, or by calling
`as_universalmotif()` to convert the data.frame back into universalmotif format,
discarding columns that cannot be stored within `universalmotif` format.
Therefore, this data structure empowers users to import and analyze motif lists
from a diverse set of sources for use with `memes` or for simple manipulation of
motif metadata using common R tools. In this way, memes enables large-scale
motif database manipulation & analysis within the preexisting R framework.

### Support for genomic range-based data

Motif analysis is often employed in ChIP-seq analysis, in which data is stored
as genomic coordinates rather than sequence. However, MEME Suite tools are
designed to work with sequences. While existing tools such as bedtools can
extract DNA sequence from gneomic coordinates, some MEME tools require fasta
headers to be specifically formatted. As a result, users must write custom code
to extract the DNA sequence for their genomic ranges of interest.

The `memes` function `get_sequence` automates extraction of
DNA sequence from genomic coordinates while simultaneously producing MEME Suite
formatted fasta headers. `get_sequence` accepts genomic-range based inputs in
`GenomicRanges` format, the *de-facto* standard for genomic coordinate
representation in R \cite{genomicranges}. Other common genomic coordinate
representations, such as bed format, are easily imported as `GenomicRanges`
objects into R using preexisting import functions, meaning `memes` users do not
have to write any custom import functions to work with range-based data using
`memes`. Sequences are returned in `Biostrings` format and can therefore be
used as inputs to all `memes` commands or as inputs to other R/Bioconductor
functions for sequence analysis.

### Data-aware motif analysis

Discriminative (or "differential") motif analysis, in which motifs are discovered in one set of
sequences relative to another set, can uncover biologically relevant motifs
associated with membership to distinct categories. For example, during analysis
of multiomic data, users can identify different functional categories of genomic
regions through integration with orthogonal datasets (e.g. categorizing
transcription factor binding sites by the presence or absence of another
factor). Although the MEME Suite allows differential enrichment testing, it does
not inherently provide a mechanism for analyzing groups of sequences in
parallel, or for performing motif analysis with any understanding of data
categories. The `memes` framework enables "data-aware" motif analysis workflows
by allowing named lists of `Biostrings` objects as input to each function. If
using `GenomicRanges`, users can split peak data on a metadata column using the
base R `split` function, and then use this result as input to `get_sequence`, which
will produce a list of `BioStrings` objects where each entry is named after the
data categories from the split column. When a list is used as input to a `memes`
function, it runs the corresponding MEME Suite tool for each object in the list.
Users can also pass the name(s) of a category to the `control` argument to
enable differential analysis of the remaining list members against the
control category sequences. In this manner, `memes` enables complex data-aware
differential motif analysis workflows using simple syntax to extend the
capabilities of the MEME Suite.

## Data visualization
The MEME Suite provides a small set of data visualizations that have limited
customizability. `memes` leverages the advantages of the R graphics environment
to provide a wide range of data visualization options that are highly
customizable. We describe two scenarios below.

The TomTom tool allows users to compare unknown motifs to a set of known motifs
to identify the best match. Visual inspection of motif comparison data is a key
step in assessing the quality and accuracy of a match. The `view_tomtom_hits`
function allows users to compare query motifs with the list of potential matches
as assigned by TomTom, similar to the commandline behavior. The
`force_best_match` function allows users to reassign the TomTom best match
to a lower-ranked (but still statistically significant) match in order to
highlight motifs with greater biological relevance (e.g. to skip over a
transcription factor that is not expressed in the experimental sample, or when
two motifs are matched equally well).

The AME tool searches for enrichment of motifs within a set of experimental
sequences relative to a control set. The meaningful result from this tool is the
statistical parameter (for example, a p-value) associated with the significance
of motif enrichment. However, AME does not provide a mechanism for visualizing
these results.

The `ame_plot_heatmap` function in `memes` returns a `ggplot2` formatted heatmap of
statistical significance of motif enrichment. If AME is used to examine motif
content of multiple groups of sequences, the `ame_plot_heatmap` function can also
return a plot comparing motif significance within multiple groups. Several
options exist to customize the heatmap values in order to capture different aspects
of the output data. The `ame_compare_heatmap_methods()` function enables users
to compare the distribution of values between samples in order to select a
threshold that accurately captures the dynamic range of their results.

## Containerized analysis maximizes availability and facilitates reproducibility
`memes` relies on a locally installed version of the MEME Suite which is
accessed by the user's R process. Although R is available on Windows, Mac, and
Linux operating systems, the MEME Suite is incompatible with Windows, limiting
its adoption by Windows users. Additional barriers also exist to installing a
local copy of the MEME Suite on compatible systems, for example, the MEME Suite
relies on several system-level dependencies whose installation can be difficult
for novice users. Finally, some tools in the MEME Suite use python to generate
shuffled control sequences for analysis, which presents a reproducibility issue
as the random number generation algorithm changed between python2.7 and python3. The
MEME Suite will build and install on both python2.7 and python3 systems,
therefore without careful consideration of this behavior, the same code run on
two systems may not produce identical results, even if using the same major
version of the MEME Suite. In order to increase access to the MEME Suite on
unsupported operating systems, and to facilitate reproducible motif analysis
using `memes`, we have also developed docker and singularity (**TODO: NOTE:**
singularity is broken for `rocker/rstudio`, so maybe omit this until fixed)
containers containing a preinstalled version of the MEME Suite along with an
R/Bioconductor analysis environment including the most recent version of `memes`
and its dependencies.  As new container versions are released, they are
version-tagged and stored to ensure reproducibility while allowing updates to
the container.

# Results

Here we briefly highlight each of `memes` current features for analyzing
ChIP-seq data. Additional detailed walkthroughs for each supported MEME Suite
tool, and a worked example using `memes` to analyze ChIP-seq data can be found
in the `memes` vignettes and the package website
([snystrom.github.io/memes](snystrom.github.io/memes)). In the following
example, we reanalyze recent work examining the causal relationship between the
binding of the transcription factor E93 to changes in chromatin accessibility
during *Drosophila* wing development \cite{Nystrom, Niederhuber 2020}. Here, we
utilize ChIP-seq peak calls for E93 which have been annotated according to the
change in chromatin accessibility observed before & after E93 binding. These
data are an emblematic example of range-based genomic data (E93 ChIP peaks)
containing additional groupings (the chromatin accessibility response following
DNA binding) whose membership may be influenced by differential motif content.
We show how `memes` syntax allows complex analysis designs, how `memes`
utilities enable deep interrogation of results, and how `memes` flexible data
structures empower users to integrate the `memes` workflow with tools offered by
other R/Bioconductor packages.

```{r, echo = F}
drake::loadd(chip_results)
```

The aforementioned ChIP-seq peaks are stored as a GRanges object with a metadata
column (`e93_chromatin_response`) indicating whether chromatin accessibility
tends to increase ("Increasing"), decrease ("Decreasing"), or remain unchanged
("Static") following E93 binding.
```{r}
head(chip_results, 3)
```

Using the `get_sequence` function, GRanges objects are converted into
DNAStringSet outputs.
```{r}
dm.genome <- BSgenome.Dmelanogaster.UCSC.dm3::BSgenome.Dmelanogaster.UCSC.dm3

all_sequences <- chip_results %>% 
  get_sequence(dm.genome)
```

```{r}
head(all_sequences)
```
In order to perform analysis within different groups of peaks, the GRanges object can be `split` on a metadata column before input to `get_sequence`.
```{r}
sequences_by_response <- chip_results %>% 
  split(mcols(.)$e93_chromatin_response) %>% 
  get_sequence(dm.genome)

```

This produces in a BStringSetList where list members contain a DNAStringSet for each group of sequences.
```{r}
head(sequences_by_response)
```

## Denovo motif analysis
The DREME tool can be used to discover short, novel motifs present in a set of
input sequences relative to a control set. The `runDreme` command is the `memes`
interface to the DREME tool. `runDreme` syntax enables users to produce complex
discriminative analysis designs using intuitive syntax. Examples of possible
designs and their syntax are compared below.

```{r, eval=F}
# use all sequences vs shuffled background
runDreme(all_sequences, control = "shuffle")

# for each response category, discover motifs against a shuffled background set
# Produces:
#  - Increasing vs Shuffled Increasing
#  - Decreasing vs Shuffled Decreasing
#  - Static vs Shuffled Static
runDreme(sequences_by_response, control = "shuffle")

# Search for motifs enriched in the "Increasing" peaks relative to "Decreasing" peaks
# Produces:
#  - Increasing vs Decreasing
runDreme(sequences_by_response$Increasing, control = sequences_by_response$Decreasing)

# Use the "Static" response category as the control set to discover motifs 
# enriched in each remaining category relative to static sites,
# Produces:
#  - Increasing vs Static
#  - Decreasing vs Static
runDreme(sequences_by_response, control = "Static")

# Combine the "Static" and "Increasing" sequences and use as a background set to
# discover motifs enriched in the remaining categories relative
# Produces:
#  - Decreasing vs Static+Increasing
runDreme(sequences_by_response, control = c("Static", "Increasing"))
```

In order to discover motifs associated with dynamic changes to chromatin
accessibility, we use static sites as the control set to discover motifs
enriched in both increasing and decreasing sites.

```{r, eval=F}
dreme_vs_static <- runDreme(sequences_by_response, control = "Static")
```

```{r, echo=FALSE}
# I secretly cache this to decrease build-time of the manuscript
drake::loadd(dreme_vs_static)
```

```{r, fig.height=4, fig.width=7, echo = FALSE}
dreme_fig_top <- purrr::map(dreme_vs_static, "motif") %>% 
  purrr::map(universalmotif::view_motifs) %>% 
  purrr::imap(~{
    .x +
      ggtitle(paste(.y, "Motifs")) +
      theme(plot.title = element_text(hjust = 0.5, size = 14))
  }) %>% 
  patchwork::wrap_plots(ncol = 2) 
```
```{r, echo=FALSE}

# Require that the minimum overlap for similarity metric is equal to the
# smallest possible overlap if the shortest motif was completely contained
# within the largest motif.
get_min_overlap <- function(df){
  min(df$length) / max(df$length)
}

get_motif_correlation <- function(df, method = "PCC"){
  df$motif %>% 
    universalmotif::compare_motifs(min.overlap = get_min_overlap(df), method = method)
}

plot_correlation_heatmap <- function(mat, name){

  min_color <- round(min(mat) - 0.1, digits = 1)
  colors <- circlize::colorRamp2(c(min_color, 1), c("white", "dodgerblue3"))

  rownames(mat) <- NULL
  ComplexHeatmap::Heatmap(mat,
                          col = colors, 
                          name = "Pearson Correlation", 
                          column_names_rot = 45,
                          column_title = paste(name, "Motif Similarity")
                          )

}
```

```{r, fig.height=5, fig.width=10, echo = FALSE}
dreme_fig_bottom <- dreme_vs_static %>% 
  purrr::map(get_motif_correlation) %>% 
  purrr::imap(plot_correlation_heatmap) %>% 
  # capture heatmaps as grobs
  purrr::map(~{grid::grid.grabExpr(draw(.x))}) %>% 
  patchwork::wrap_plots(nrow = 1) 
```

These data are readily visualized using the `universalmotif::view_motifs`
function. Visualization of the results reveals two distinct motifs associated
with decreasing and increasing sites \ref{fig:dec_v_inc_dreme}. In this
analysis, the *de-novo* discovered motifs within each category appear visually
similar to each other, however, the MEME Suite does not provide a mechanism to
compare groups of motifs based on all pairwise similarity metrics. We utilized
the `universalmotif::compare_motifs` function to compute Pearson correlation
coefficients for each set of motifs to quantitatively assess motif similarity
\ref{fig:dec_v_inc_dreme}.

```{r dec_v_inc_dreme, fig.height=10, fig.width=10, echo=F, fig.cap="\\label{fig:dec_v_inc_dreme}PWMs of discovered motifs in Decreasing (A) and Increasing (B) sites. C,D Pearson correlation heatmaps comparing similarity of Decreasing (C) and Increasing (D) motifs."}
patchwork::wrap_plots(dreme_fig_top, dreme_fig_bottom, ncol = 1) +
  patchwork::plot_annotation(tag_levels = "A")
```

Transcription factors often bind with sequence specificity to regulate activity
of nearby genes. Therefore, comparison of *de-novo* motifs with known DNA
binding motifs for transcription factors can be a key step in identifying
transcriptional regulators. The TomTom tool is used to compare unknown motifs
against a list of known motifs to identify matches, `memes` provides the
`runTomTom` function as an interface to this tool. By passing the results from
`runDreme` into `runTomTom` and searching within a database of *Drosophila*
transcription factor motifs, we can identify candidate transcription factors
that may bind the motifs associated with increasing and decreasing chromatin
accessibility.

```{r}
dreme_vs_static_tomtom <- runTomTom(dreme_vs_static, database = "data/flyFactorSurveyCleaned.meme", dist = "ed")
```

Using this approach, the results can be visualized using  the
`view_tomtom_hits()` function to visually inspect the matches assigned by
TomTom, providing a simple way to assess the quality of the match.  A
representative plot of these data is shown in \ref{fig:tomtom_example}, while
the full results can be viewed in \ref{tab:tomtom_table}. This analysis revealed
that the motifs associated with decreasing sites are highly similar to the E93
motif. Conversely, motifs found in increasing sites match the transcription
factors br, l(3)neo38, gl, and lola \ref{tab:tomtom_table}. These data support
the hypothesis that E93 binding to its motif may play a larger role in chromatin
closing activity, while binding of the transcription factor br is associated
with chromatin opening.

```{r tomtom_example, echo = F, fig.height=3, fig.width=7.5, fig.cap="\\label{fig:tomtom_example}Representative plots generated by view_tomtom_hits showing the top hit for motifs discovered in decreasing (A) and increasing (B) peaks."}
plot_top_hit <- function(x){
  x %>% 
    view_tomtom_hits(1)
}

top_hit_grid <- function(x, title){
  x %>% 
    patchwork::wrap_plots(nrow = 1)
}

# Plotting every hit:
# decided to just use a representative sample insted
#dreme_vs_static_tomtom %>% 
#  purrr::imap(~{plot_top_hit(.x) %>% top_hit_grid(., .y)}) %>% 
#  patchwork::wrap_plots(ncol = 1)

dreme_vs_static_tomtom %>% 
  purrr::imap(~{
    .x %>% 
      dplyr::mutate(name = paste0(.y, " ", name)) %>% 
      update_motifs() %>% 
      plot_top_hit() %>% 
      .[[2]]
    }) %>% 
  patchwork::wrap_plots() +
    patchwork::plot_annotation(tag_levels = "A")
```
```{r, echo=F}
library(gt)

dreme_vs_static_tomtom %>% 
  dplyr::bind_rows(.id = "category") %>% 
  dplyr::mutate(category = paste(category, "Motifs")) %>% 
  dplyr::arrange(category, rank) %>% 
  dplyr::select(category, name, best_match_altname) %>% 
  dplyr::rename("Best Match TF" = best_match_altname) %>% 
  gt(rowname_col = "name",
     groupname_col = "category")
     # this feature is in development branch??
     # search SO: gt table reference
     #lab = "tab:tomtom_table")

```

## Motif enrichment testing using runAme

Discovery and matching of *de-novo* motifs is only one way to find candidate
transcription factors within target sites. Indeed, in many instances, requiring
that a motif is recovered *de-novo* is not ideal, as these approaches are
"discovery based". Another approach, implemented
by the AME tool, is to search for known motif instances in target sequences and
test for their overrepresentation relative to a background set of sequences
\cite{ame}. The `runAme()` function is the `memes` interface to the AME tool. It
accepts a set of sequences as input and control sets, and will peform enrichment
testing using a provided motif database for occurrances of each provided motif.

A major limitation of this approach is that transcription factors containing
similar families of DNA binding domain often possess highly similar motifs,
making it difficult to identify the "true" binding factor associated with an
overrepresented motif. Additionally, when searching for matches against a motif
database, AME must account for multiple testing, therefore using a larger than
necessary motif database can produce a large multiple testing penalty, limiting
sensitivity of detection. One way to overcome these limitations is to limit the
transcription factor motif database to include only motifs for transcription
factors expressed in the sample of interest. Accounting for transcription factor
expression during motif analysis has been demonstrated to increase the
probability of identifying biologically relevant transcription factor candidates
\cite{hector franco's STAR paper, others?}.

The `memes` universalmotif data.frame structure can be used to integrate
expression data with a motif database to remove entries for transcription
factors that are not expressed. To do so, we import a *Drosophila*
transcription factor motif database generated by the Fly Factor Survey and
convert to universalmotif data.frame format \cite{fly factor}. In this database,
the `altname` column stores the gene symbol.

```{r}
fly_factor <- universalmotif::read_meme("data/flyFactorSurveyCleaned.meme") %>% 
  as_universalmotif_dataframe()
```

```{r, echo = FALSE}
head(fly_factor, 3)
```

Next, we import a pre-filtered list of genes expressed in a timecourse of
*Drosophila* wing development.

```{r}
wing_expressed_genes <- read.csv("data/wing_expressed_genes.csv")
```

```{r, echo = FALSE}
n_fly_factor_expressed <- fly_factor %>% 
  dplyr::filter((altname %in% wing_expressed_genes$symbol)) %>% 
  nrow()
n_fly_factor_unexpressed <- fly_factor %>% 
  dplyr::filter(!(altname %in% wing_expressed_genes$symbol)) %>% 
  nrow()
```

Finally, we subset the motif database to only expressed genes using `dplyr`
data.frame subsetting syntax (note that base R subsetting functions operate
equally well on these data structures), then convert the data.frame back into
`universalmotif` format using `as_universalmotif`. This filtering step removes
`r (n_fly_factor_unexpressed / nrow(fly_factor)) %>% {. * 100} %>% round()`% of
entries from the original database, greatly reducing the multiple-testing
correction.

```{r}
fly_factor_expressed <- fly_factor %>% 
  dplyr::filter(altname %in% wing_expressed_genes$symbol) %>% 
  as_universalmotif()
```

`runAme` syntax is identical to `runDreme` in that discriminative designs can be
constructed by calling list entries by name. Data can be visualized using
`ame_plot_heatmap()`, revealing that the E93 motif is strongly enriched at
Decreasing sites, while l(3)neo38, lola, and br are enriched in Increasing
sites, supporting the *de-novo* discovery results \ref{fig:ame_heatmap}.
Additionally, AME detects several other transcription factor motifs that
distinguish decreasing and increasing sites, providing additional clues to
potential factors that also bind with E93 affect chromatin accessibility.

```{r}
ame_vs_static <- runAme(sequences_by_response, control = "Static", database = fly_factor_expressed)
```

```{r ame_heatmap, fig.height=3, fig.width=7.5, echo = FALSE, fig.cap="\\label{fig:ame_heatmap} A heatmap of the -log10(adj.pvalues) from AME enrichment testing within increasing and decreasing sites."}
ame_vs_static %>% 
  dplyr::bind_rows(.id = "category") %>% 
  dplyr::mutate(category = paste0(category, "\nAccessibility")) %>% 
  dplyr::group_by(category, motif_alt_id) %>% 
  dplyr::filter(adj.pvalue == min(adj.pvalue)) %>% 
  ame_plot_heatmap(group = category, id = motif_alt_id) +
    labs(x = "Transcription Factor",
         y = "ChIP Peak Type")
```

### Motif matching using runFimo

A striking result from these analyses is that the E93 motif is so strongly
enriched within E93 ChIP peaks that decrease in accessibility. Significance of
motif enrichment can be driven by several factors, such as quality of the query
motif relative to the canonical motif or differences in motif number in one
group relative to other groups. These questions can be explored directly by
identifying motif occurrances in target regions and examining their properties.
FIMO allows users to match motifs in input sequences while returning information
about the quality of each match in the form of a quantitative score \cite{fimo}.

In order to examine the properties of the E93 motif between different ChIP
peaks, we scan all E93 ChIP peaks for matches to the Fly Factor Survey E93 motif
using `runFimo`. Results are returned as a GRanges object containing the
positions of each motif. Setting `skip_matched_sequence` to `FALSE` adds the
sequence of the detected motif to the output.

```{r}
e93_fimo <- runFimo(all_sequences, fly_factor_expressed$Eip93F_SANGER_10, skip_matched_sequence = FALSE, thresh = 1e-3)
```

Using `plyranges`, matched motifs can be joined with the metadata of the ChIP peaks with which they intersect.

```{r}
e93_fimo %<>% 
  plyranges::join_overlap_intersect(chip_results)
```

```{r, echo=F}
e93_stack <- e93_fimo %>% 
  data.frame %>%  
  split(.$e93_chromatin_response) %>% 
  purrr::imap(~{
    universalmotif::create_motif(.x$matched_sequence, name = paste(.y, "E93"))
  }) %>% 
  rev %>% 
  {
    m <- .
    e93_flyfactor <- fly_factor_expressed$Eip93F_SANGER_10
    e93_flyfactor@name <- "E93 FlyFactorSurvey"

    m$e93 <- e93_flyfactor
    return(m)
  } %>% 
  rev %>% 
  universalmotif::view_motifs()
```

```{r, echo=F}
n_motifs <- chip_results %>% 
  plyranges::mutate(has_match = ifelse(. %over% e93_fimo, "Match", "No Match")) %>% 
  data.frame %>% 
  ggplot(aes(e93_chromatin_response)) +
    geom_bar(aes(fill = forcats::fct_rev(has_match)), position = "fill") +
     scale_fill_manual(values = c("Match" = "firebrick",
                                  "No Match" = "Black")) +
     labs(fill = "E93 Motif Match",
          y = "Fraction of Sites",
          x = "Response to E93 binding") 
```

```{r, fig.height=5, fig.width=3, echo=F}
motif_quality <- e93_fimo %>% 
  data.frame %>% 
  ggplot(aes(e93_chromatin_response, score)) +
    geom_boxplot(aes(fill = e93_chromatin_response), notch = TRUE) +
    theme_bw() +
    labs(y = "FIMO Score",
         x = "Chromatin Response in E93 ChIP Peak") +
    theme(legend.position = "none")
```

Using this approach we can deeply examine the properties of the E93 motif within
each chromatin response category. First, by counting the number of E93 motifs
within each category, we demonstrate that Decreasing sites are more likely than
increasing or static sites to contain an E93 motif \ref{fig:fimo}A. We extend
these observations by rederiving position-weight matrices from sequences
matching the E93 motif within each category, allowing visual inspection of motif
quality across groups \ref{fig:fimo}B. Notably, differences in quality at base
positions 8-12 appear to distinguish increasing from decreasing motifs, where
decreasing motifs are more likely to have strong A bases at positions 8 and 9,
while E93 motifs from increasing sites are more likely to have a T base pair at
position 12 \ref{fig:fimo}B. Examination of bulk FIMO scores (where higher
scores represent motifs more similar to the reference) also reveals differences
in E93 motif quality between groups, in particular, E93 motifs from Decreasing
sites have higher scores \ref{fig:fimo}C. Together, these data demonstrate that
a key distinguishing factor between whether a sites will decrease or increase in
chromatin accessibility following E93 binding is the number and quality of E93
motifs at that site.

In summary, `memes` establishes a powerful motif analysis environment by
leveraging the speed and utility of the MEME Suite set of tools in conjunction
with the flexible and extensive R/Bioconductor package landscape.

```{r fimo, fig.width=7.5, fig.height=9, fig.cap="\\label{fig:fimo}A. Stacked barplot showing fraction of each chromatin response category containing E93 motif matches. B. PWMs generated from E93 motif sequences detected in each chromatin response category. C. Boxplot of FIMO Score for each E93 motif within each chromatin response category. Outliers are plotted as distinct points."}
#patchwork::wrap_plots(n_motifs, motif_quality, patchwork::guide_area(), e93_stack, widths = c(0.2, 0.3, 0.5))
patchwork::wrap_plots(n_motifs, patchwork::guide_area(), 
                      e93_stack, motif_quality, 
                      ncol = 2, widths = c(0.6, 0.4),
                      guides = 'collect') +
    patchwork::plot_annotation(tag_levels = "A")
```

# Availability and Future developments:

The `memes` package source code and installation instructions are available on
github: [github.com/snystrom/memes](github.com/snystrom/memes). Documentation is
stored in the package vignettes, and also available at the package website:
[snystrom.github.io/memes](snystrom.github.io/memes). The `memes_docker`
containers are available from DockerHub
[https://hub.docker.com/r/snystrom/memes_docker](https://hub.docker.com/r/snystrom/memes_docker),
and their source code is hosted at github:
[https://github.com/snystrom/memes_docker](https://github.com/snystrom/memes_docker).

This manuscript was automatically generated using Rmarkdown within the
`memes_docker` container (TODO: ADD VERSION). Its source code, raw data, and
instructions to reproduce all analysis can be found at
[github.com/snystrom/memes_paper/](github.com/snystrom/memes_paper/).

In the future we hope to add additional data visualizations for examining motif
positioning within features. We will continue to add support for additional MEME Suite
tools in future versions of the package. Finally, we hope to improve the `memes`
tooling for analyzing amino-acid motifs, which although fully supported by our
current framework, may require extra tools which we have not considered.

# Acknowledgements
We would like to acknowledge Megan Justice, and members of the McKay lab for
helpful comments and feedback.